Transferability of clinical prediction models for early trauma care

### 2019-08-25 ManuscriptMarkdown.Rmd
Add change in how to model vital signs to manuscript text and add reference

## RESULTS
  ## Total patients and NA
  total.number.of.patients <- nrow(results$data.sets.before.imputations$multi.centre.vs.single.centre$multi.centre)
  total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
  n.missing.data.time.of.trauma <- Results$n.missing.date.time.of.trauma
  ## Sample NA
  na.vector <- unlist(results$NA.info.sample)
  names(na.vector) <- c("high volume", "high volume", "low volume", "low volume", "metropolitan", "metropolitan", "non metropolitan", "non metropolitan", "multi centre", "multi centre", "single centre", "single centre")
  highest.percent.missing.number <- max(na.vector[na.vector<100])
  highest.percent.missing.name <- names((na.vector[na.vector == highest.percent.missing.number]))
  ## Variable NA
  number.variable.na <- max(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na)
  percent.variable.na <- results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$percentage[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na]
  name.variable.na <- row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,])
  gcs.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_gcs_sum",])
  sbp.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_sbp_value",])
  ## Development and validation
  ## Imputation data
  high.low.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "high.volume"]
  metro.non.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "metropolitan"]
  multi.single.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "multi.centre"]
  max.imp <- max(results$original.stats$Total.imputations)
  ## Validation data
  sorted.validation.mistriage <- sort(results.all.data[row.names(results.all.data) == "Validation.mistriage",], decreasing = FALSE)
  lowest.val.mistriage <- sorted.validation.mistriage[1]
  val.undertriage <- results.all.data["Validation.undertriage", colnames(lowest.val.mistriage)]
  val.overtriage <- results.all.data["Validation.overtriage", colnames(lowest.val.mistriage)]
  highest.val.mistriage <- sorted.validation.mistriage[6]
  ## Comparison
  ## Transfer data
  sorted.transfer.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transfer.mistriage",], decreasing = FALSE)
  lowest.tran.mistriage <- sorted.transfer.mistriage[1]
  highest.tran.mistriage <- sorted.transfer.mistriage[6]
  ## Comparison data
  sorted.comp.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transferred.mistriage.minus.buddy.local.mistriage",], decreasing = FALSE)
  highest.comp.mistriage <- sorted.comp.mistriage[6]
  worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]

@@

## Quantitative variables
GCS was modelled as a continuous linear term, and SBP and RR were modelled using 
restricted cubic splines with four knots whenever possible, placed at equally spaced 
percentiles. When describing the participant characteristics, quantitative variables 
were presented as continuous. ISS was presented as dichotomous, using ISS > 15 as the cutoff.

@@

### Development and validation sample
The samples in each data set were split into
two subsamples using a temporal split based on the date of trauma. The earlier
subsample was the development sample, and the later subsample was the validation
sample. The development sample contained 70 events (events being patients who
died within 30 days of the trauma) and all non-events (non-events being patients who
survived 30 days past the trauma) during the same time. The rationale for
including 70 events is that at least 10 events per free parameter were needed in
the logistic regression to obtain stable coefficient estimates [@RN30]. The
validation sample contained 100 events and at least 100 non-events,
which was suggested as the minimum number by Vergouwe in 2005 for external
validation samples [@RN37]. Figure 1 shows the development of the samples using 
the high- and low-volume centre data set as an example. The minimum sample sizes 
of the development and validation samples were 140 and 200, respectively. 
We performed analyses only on data sets for which all samples included at 
least the minimum sample size.

@@

### Missing data
We used multiple imputation using chained equations (mice) to handle missing data [@RN41]. 
The number of imputations created for each data set was equal to the highest 
percentage of missing data in that data set. Quantitative variables were imputed 
using predictive mean matching, and qualitative variables were imputed using logistic 
regression. SBP and RR were transformed as restricted cubic splines before imputation. 
All analyses outlined above were then conducted separately in
each imputed dataset. We present the combined results as the median point
estimate across imputations along with an empirical bootstrap of the 25th and
75th percentiles across imputations, i.e. the lower bound of the presented
interval is the lower bound of a 95% CI of the 25th percentile, and the upper
bound is the upper bound of a 95% CI of the 75th percentile. This combined CI
was referred to as an uncertainty interval (UI) and was used to express the
added uncertainty associated with the imputation procedure to handle missing
data; as such, it is more conservative than a standard 95% CI.

>>>>

## RESULTS
  ## Total patients and NA
  total.number.of.patients <- nrow(results$data.sets.before.imputations$multi.centre.vs.single.centre$multi.centre)
  total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
  n.missing.data.time.of.trauma <- results$n.missing.date.time.of.trauma
  ## Sample NA
  na.vector <- unlist(results$NA.info.sample)
  names(na.vector) <- c("high volume", "high volume", "low volume", "low volume", "metropolitan", "metropolitan", "non metropolitan", "non metropolitan", "multi centre", "multi centre", "single centre", "single centre")
  highest.percent.missing.number <- max(na.vector[na.vector<100])
  highest.percent.missing.name <- names((na.vector[na.vector == highest.percent.missing.number]))
  ## Variable NA
  number.variable.na <- max(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na)
  percent.variable.na <- results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$percentage[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na]
  name.variable.na <- row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,])
  gcs.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_gcs_sum",])
  sbp.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_sbp_value",])
  ## Development and validation
  ## Imputation data
  high.low.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "high.volume"]
  metro.non.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "metropolitan"]
  multi.single.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "multi.centre"]
  max.imp <- max(results$original.stats$Total.imputations)
  ## Validation data
  sorted.validation.mistriage <- sort(results.all.data[row.names(results.all.data) == "Validation.mistriage",], decreasing = FALSE)
  lowest.val.mistriage <- sorted.validation.mistriage[1]
  val.undertriage <- results.all.data["Validation.undertriage", colnames(lowest.val.mistriage)]
  val.overtriage <- results.all.data["Validation.overtriage", colnames(lowest.val.mistriage)]
  highest.val.mistriage <- sorted.validation.mistriage[6]
  ## Comparison
  ## Transfer data
  sorted.transfer.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transfer.mistriage",], decreasing = FALSE)
  lowest.tran.mistriage <- sorted.transfer.mistriage[1]
  highest.tran.mistriage <- sorted.transfer.mistriage[6]
  ## Comparison data
  sorted.comp.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transferred.mistriage.minus.buddy.local.mistriage",], decreasing = FALSE)
  highest.comp.mistriage <- sorted.comp.mistriage[6]
  worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]

@@

## Quantitative variables
GCS, SBP, and RR were modelled as continuous linear terms. In the original study
plan it was specified that SBP and RR would be modelled using restricted cubic
splines with four knots. This approach however resulted in an unstable
imputation model and we therefore decided to simplify the analysis model. ISS
was presented as dichotomous, using ISS > 15 as the cutoff.

@@

### Development and validation sample
The samples in each data set were split into two subsamples using a temporal
split based on the date of trauma. The earlier subsample was the development
sample, and the later subsample was the validation sample. The development
sample contained 75 events (events being patients who died within 30 days of the
trauma) and all non-events (non-events being patients who survived 30 days past
the trauma) during the same time. The rationale for including 75 events was that
25 events per variable have been shown to result in less biased coefficient
estimates compared to the more standard 10 events per variable [@RN30;
@Courvoisier2011]. The validation sample contained 100 events and at least 100
non-events, which was suggested as the minimum number by Vergouwe in 2005 for
external validation samples [@RN37]. Figure 1 shows the development of the
samples using the high- and low-volume centre data set as an example. The
minimum sample sizes of the development and validation samples were 140 and 200,
respectively.  We performed analyses only on data sets for which all samples
included at least the minimum sample size.

@@

### Missing data
We used multiple imputation using chained equations (mice) to handle missing
data [@RN41].  The number of imputations created for each data set was equal to
the highest percentage of missing data in that data set. Quantitative variables
were imputed using predictive mean matching, and qualitative variables were
imputed using logistic regression.  All analyses outlined above were then
conducted separately in each imputed dataset. We present the combined results as
the median point estimate across imputations along with an empirical bootstrap
of the 25th and 75th percentiles across imputations, i.e. the lower bound of the
presented interval is the lower bound of a 95% CI of the 25th percentile, and
the upper bound is the upper bound of a 95% CI of the 75th percentile. This
combined CI was referred to as an uncertainty interval (UI) and was used to
express the added uncertainty associated with the imputation procedure to handle
missing data; as such, it is more conservative than a standard 95% CI.
